Apparent preferences for cognitive effort fade when multiple forms of effort and delay are interleaved in a foraging environment

### 2020-10-26 Cost_manuscript.Rmd
more intro

Outline:
- Effort and delay tend to be aversive, and neuroeconhas traditionally examined them as such
- But there is a flipside of effort: it can be attractive
- One issue is that even for studies that compare across demands, they do so by quantifying discounting curves
- A less explored avenue that might get at this though is the comparison among demands
- Ecological foraging paradigms can help us with this approach, allowing for people to search for their preferred demands
- In addition, foraging rests on a rich modeling history
- We can model whether preferences for demands are due to either 1) modulating cost or embuing value relative to optimal behavior; 2) whether people attend to focal decision elements or global environmental attributes when they evaluate demanding offers
- In this study....

  Prospectus: Evaluating whether to pursue a desired outcome often requires assessing the demands required for obtaining it. In psychology and economics, demands for both time and effort have traditionally been identified as costs that detract from the net subjective value of rewards (Westbrook & Braver, 2015). Indeed, decision makers tend to prefer low-effort and immediate rewards over larger rewards that demand longer delays and higher effort (Ainslie, 1975; Hull, 1943; Kool & Botvinick, 2018; Kool, McGuire, Rosen, & Botvinick, 2010; Treadway, Buckholtz, Schwartzman, Lambert, & Zald, 2009; Westbrook, Kester, & Braver, 2013). However, people often change their attitudes towards exerting effort in everyday decisions. For example, we might choose to take the effortful route home that we normally avoid rather than waiting for a delayed train. . Understanding the sporadic attractiveness of effort exemplified above has been elusive, and presents an important research avenue (Inzlicht, Shenhav, & Olivola, 2017). (add Loewenstein 1987 and other explanatory candidates here. I think the goal will be that there are theories speculating about when demands can be valuable for a single choice decision, or in TAFC, but point out inconclusiveness. The next paragraph should say that what's missing is comparing among costs)
  A key limitation in understanding this phenomenon is that studies often compare the reduction of subjective value imposed by each type of demand, often through discounting functions that capture how much less appealing a reward becomes as a function of the amount or intensity of its associated behavior [1,3,5,8,9,14]. Although this line of work suggests that discounting functions are similar for delay and effort [8,15,16], there is ongoing debate about the shape of the discounting function across effort types [2,8,11,17,18], and discount rates do not always correlate across cost domains [9,16]. Furthermore, this approach fails to interrogate which type of demand people would select. Flexible preferences for effort might stem from the availability of other types of demands in the same choice environment (Kool & Botvinick, 2018).
  Single choice foraging paradigms provide a natural way to probe preferences among demands, by asking individuals whether the current offer outweighs what could be obtained and experienced during that time instead (Mobbs, Trimmer, Blumstein, & Dayan, 2018). In addition, foraging tasks can allow individuals to emphasize different elements of the decision context, and rest on a rich modeling history that can help to computationally capture such emphases (Stephens & Krebs, 2019).
  In this study, we will test the hypothesis that preferences will vary depending on whether individuals face a single or multiple types of behavioral demands, and identify what factors drive these choices. Based on two foraging experiments and computational modeling, we will show that 1) preferences for mental effort are stable when individuals face a single form of demand, but fade over time when other costs are also experienced; and 2) the phenomenon can be explained by a model in which preferences depend on how participants construe the amount of time a task takes, which could hypothetically be recalibrated when all tasks are interleaved.
  In this study, we will behaviorally and computationally test the hypothesis that the attractiveness of a reward that requires effort or delay depends on whether rewards associated with alternative demands are available in the same context. Using a foraging paradigm will enable us to 1) assess whether people integrate the statistics of the environment to generate strategies that maximize their earnings; 2) compare how the presence of effort or delay shifts such strategies, as reflected by acceptance rates; 3) examine if acceptance rates are further altered by foraging for costs in addition to rewards; and 4) test a combination of foraging (Constantino & Daw, 2015; Stephens & Krebs, 2019; Fawcett, McNamara, & Houston, 2012) to examine whether preferences for demands reflect a subjective reduction of cost or boost of value, and determine any factors besides contextual manipulations that participants consider in order to adjust their choices.

@@

- Make sure to note that we controlled for time as a confound for effort, as more effortful tasks normally take longer
- What we think: subjective timing is shortened by the way we implemented cognitive effort. This can be recalibrated if many task types are experienced. Maybe the implementation allowed for counting. Alternatively, working memory engagement could produce timing issues (Buhi n Gershman). Also look at what Andrew Westbrook shared. Add this to a potential fatigue, although that would also mean that waiting would be accepted less over time in exp 1.
- Things that don't seem to be at play here, but could still play a general role in mental effort preferences: 1) A demand’s cost is not fully intrinsic (within vs between differences); 2) Effort preferences unlikely just due to learned industriousness or efficacy as amitais lab sees it (differences between effort types, and everyone met a good performance level per training); 3) avoiding boredom (global decline in exp 2, not a preference exchange). 4) Literally counting tasks, since mental effort was discretely presented but everything else was contiguous. If that were the case, dividing the handling time by 2 would work, but not sure if it would translated into exp 2. 5) fatigue: there's a similar gradual decrease in acceptances in both experiments, except for the wait condition in exp 1. Fatigue would apply equally to both, or alternatively it might not bring all down to the same level (I guess the point could be made, but it's not that people who complete more at the beginning decrease acceptances more, as one would expect from fatigued people)
- Deviations in the willingness to pursue effort can predict depression relapse (https://jamanetwork.com/journals/jamapsychiatry/article-abstract/2761562), and motivational factors affecting effort aversion can be heterogenous between depression and schizophrenia (https://www.researchgate.net/publication/337361164_Pathways_to_Motivational_Impairments_in_Psychopathology_Common_Versus_Unique_Elements_Across_Domains). Like, the aversion is due to focusing on different elements of the task.
- Talk about future experiments: if this is subjective time, changing the pacing of the tasks should change the acceptance rates for cognitive effort
- It's true that we didnt calibrate cog effort per subject. But instead we brought all subjects to the same level of performance, and we saw that as time progressed they became less likely to accept longer periods of effort.

>>>>

Outline:
- Effort and delay tend to be aversive, and neuroeconhas traditionally examines them as such
- But there is a flipside of effort: it can be attractive
- One issue is that even for studies that compare across demands, they do so by quantifying discounting curves
- A less explored avenue that might get at this though is the comparison among demands
- Ecological foraging paradigms can help us with this approach, allowing for people to search for their preferred demands
- In addition, foraging rests on a rich modeling history
- We can model whether preferences for demands are due to either 1) modulating cost or embuing value relative to optimal behavior; 2) whether people attend to focal decision elements or global environmental attributes when they evaluate demanding offers
- In this study....

  Evaluating whether to pursue a desired outcome often requires assessing the demands required for obtaining it. The intuitive reaction to this evaluation is aversion. In psychology and economics, demands for both time and effort have traditionally been identified as costs that detract from the net subjective value of rewards (Westbrook & Braver, 2015). Indeed, decision makers tend to prefer low-effort and immediate rewards over larger rewards that demand longer delays and higher effort (Ainslie, 1975; Hull, 1943; Kool & Botvinick, 2018; Kool, McGuire, Rosen, & Botvinick, 2010; Treadway, Buckholtz, Schwartzman, Lambert, & Zald, 2009; Westbrook, Kester, & Braver, 2013). However, people often change their attitudes towards exerting effort in everyday decisions, resulting in pursual of effortful avenues instead (Inzlicht, Shenhav, & Olivola, 2017). Much experimental and theoretical work surrounding the sporadic attractiveness of effort suggests a reciprocal relationship between effort and outcome, whereby exerting effort boosts the perceived value of its outcomes (kacelnik), and positive outcomes imbue value into their preceding effortful behavior (learned industriousness cite). Even though the prevalent emphasis on these retrospective value associations can help explain a variety of scenarios, it is still in its infancy (be more explicit).

  What is often missing from these accounts is the common case of choosing among behavioral demands, which is partly due to the scope of experimental designs. Studies tend to compare the reduction of subjective value imposed by each type of demand, often through discounting functions that capture how much less appealing a reward becomes as a function of the amount or intensity of its associated behavior [1,3,5,8,9,14]. Although this line of work suggests that discounting functions are similar for delay and effort [8,15,16], there is ongoing debate about the shape of the discounting function across effort types [2,8,11,17,18], and discount rates do not always correlate across cost domains [9,16]. Rather than arising from changes in the decay of value, flexible preferences for effort might stem from the availability of other types of demands in the same choice environment—in other words, from its context (Kool & Botvinick, 2018). For example, exerting high effort for a reward might be aversive, but the same level of work might become valuable when compared to the boredom of passive waiting. Now add charitable giving evidence from inzlicht paper.

  Single choice foraging paradigms provide a natural way to probe preferences among demands, by asking individuals whether the current offer outweighs what could be obtained and experienced during that time instead (Mobbs, Trimmer, Blumstein, & Dayan, 2018). In addition, foraging tasks can allow individuals to emphasize different elements of the decision context, and rest on a rich modeling history that can help to computationally capture such emphases (Stephens & Krebs, 2019). Foraging has attracted high scientific interest due to its ecological validity, which is rooted in evolutionary choice behaviors [49]. The paradigm requires the decision maker to choose whether to engage with a single presented offer or skip it in search of a better one. A key advantage of foraging paradigms is that people can rapidly adopt stable, rewardmaximizing strategies [50]. Add that Fawcett bit…

  In this study, we will test the hypothesis that preferences will vary depending on whether individuals face a single or multiple types of behavioral demands. Leveraging the ecological and computational richness of foraging paradigms will provide new insights on a number of open questions. First, the existence of a well-defined optimal choice strategy in foraging paradigms is useful for identifying when demands boost the value of rewards (i.e. overharvesting poor offers). Second, increasing the level of exposure to types of demands over two experiments will let us probe whether their cost is intrinsic or relative. Third, confronting participants with both mental and physical effort will let us probe whether these preferences are domain general or task-specific. Finally, modeling will help us disambiguate what drives these preferences. We predict that..

####-------######
  Single choice foraging paradigms provide a natural way to probe preferences among demands, by asking individuals whether the current offer outweighs what could be obtained and experienced during that time instead (Mobbs, Trimmer, Blumstein, & Dayan, 2018). In addition, foraging tasks can allow individuals to emphasize different elements of the decision context, and rest on a rich modeling history that can help to computationally capture such emphases (Stephens & Krebs, 2019).
  In this study, we will test the hypothesis that preferences will vary depending on whether individuals face a single or multiple types of behavioral demands, and identify what factors drive these choices. Based on two foraging experiments and computational modeling, we will show that 1) preferences for mental effort are stable when individuals face a single form of demand, but fade over time when other costs are also experienced; and 2) the phenomenon can be explained by a model in which preferences depend on how participants construe the amount of time a task takes, which could hypothetically be recalibrated when all tasks are interleaved.
  In this study, we will behaviorally and computationally test the hypothesis that the attractiveness of a reward that requires effort or delay depends on whether rewards associated with alternative demands are available in the same context. Using a foraging paradigm will enable us to 1) assess whether people integrate the statistics of the environment to generate strategies that maximize their earnings; 2) compare how the presence of effort or delay shifts such strategies, as reflected by acceptance rates; 3) examine if acceptance rates are further altered by foraging for costs in addition to rewards; and 4) test a combination of foraging (Constantino & Daw, 2015; Stephens & Krebs, 2019; Fawcett, McNamara, & Houston, 2012) to examine whether preferences for demands reflect a subjective reduction of cost or boost of value, and determine any factors besides contextual manipulations that participants consider in order to adjust their choices.

studies have not focused on how effort and delay might interact in the common case of choosing among demands.

@@

  Understanding what drives the sporadic attractiveness of effortful avenues remains elusive. In this study, we...1) preferences for mental effort are stable when individuals face a single form of demand, but fade over time when other costs are also experienced; and 2) the phenomenon can be explained by a model in which preferences depend on how participants construe the amount of time a task takes, which could hypothetically be recalibrated when all tasks are interleaved.

- Make sure to note that we controlled for time as a confound for effort, as more effortful tasks normally take longer
- What we think: subjective timing is shortened by the way we implemented cognitive effort. This can be recalibrated if many task types are experienced. Maybe the implementation allowed for counting. Alternatively, working memory engagement could produce timing issues (Buhi n Gershman). Also look at what Andrew Westbrook shared. Add this to a potential fatigue, although that would also mean that waiting would be accepted less over time in exp 1.
- Things that don't seem to be at play here, but could still play a general role in mental effort preferences: 1) A demand’s cost is not fully intrinsic (within vs between differences); 2) Effort preferences unlikely just due to learned industriousness or efficacy as amitais lab sees it (differences between effort types, and everyone met a good performance level per training); 3) avoiding boredom (global decline in exp 2, not a preference exchange). 4) Literally counting tasks, since mental effort was discretely presented but everything else was contiguous. If that were the case, dividing the handling time by 2 would work, but not sure if it would translated into exp 2. 5) fatigue: there's a similar gradual decrease in acceptances in both experiments, except for the wait condition in exp 1. Fatigue would apply equally to both, or alternatively it might not bring all down to the same level (I guess the point could be made, but it's not that people who complete more at the beginning decrease acceptances more, as one would expect from fatigued people). Also worth mentioning that even if this is due to need for cognition (i.e. Ceyda; Laura; Cacioppo et al 1996), the collapsing in the second experiment shows that preferences for cognitive effort cannot be just that.
- Deviations in the willingness to pursue effort can predict depression relapse (https://jamanetwork.com/journals/jamapsychiatry/article-abstract/2761562), and motivational factors affecting effort aversion can be heterogenous between depression and schizophrenia (https://www.researchgate.net/publication/337361164_Pathways_to_Motivational_Impairments_in_Psychopathology_Common_Versus_Unique_Elements_Across_Domains). Like, the aversion is due to focusing on different elements of the task.
- Talk about future experiments: if this is subjective time, changing the pacing of the tasks should change the acceptance rates for cognitive effort
- It's true that we didnt calibrate cog effort per subject. But instead we brought all subjects to the same level of performance, and we saw that as time progressed they became less likely to accept longer periods of effort.
